WORKING TITLE: The adaptive value of green nest material across birds: A systematic review and meta-analysis

### 2025-03-17 docs/manuscript_GNM.qmd
Added a little more to methods

pacman::p_load(readr)

@@

@@

To obtain any missing data and to locate any grey literature, we wrote to the corresponding author of the articles that passed fulltext screening using a standardized email template (\@figsupp-email). We obtained one unpublished dataset ... \[calculate estimates here\].. Another unpublished dataset was provided by a co-author of this paper. We extracted data for `{r}  nrow(selected_abstract_screening_repeat) + (nrow(selected_abstract_screening)-nrow(excluded_fulltext))` articles and 2 unpublished studies.

<!--# Fix the code for generating this number later by using the data file directly to calculate the number of article -->

@@

We calculated two effect sizes for each pairwise comparision-log response ratio (lnRR) and standardised mean difference with heteroscedasticity correction (SMDH), using escalc function (measure= ROM and SMDH respectively, vtype= "LS") in metafor package (@metafor). We specified treatment group as the numerator and the control group as the denominator for lnRR and treatment minus control for SMDH, such that the positive values indicate the trait value increased in the treatment group. In cases such as measures of parasite load or prevalence, laying or hatching dates, courtship time, scab score etc., the sign was inverted to ensure comparability across all fitness proxies. In all cases, the effect size were coded such that the fitness proxies were positively correlated with the overall fitness.

For studies reporting only inferential statistics (e.g., t-tests, ANCOVA, or contingency tables), we estimated SMD values using the equations provided by @thehand2019. lnRR cannot be computed from inferential statistics, and only SMD under the assumption of homoskedasticity was derived from such data. Since lnRR can only be used for ratio scale data, we excluded all effect sizes with a negative value (n=?) when calculating lnRR.

The sample sizes used was the number of nests in each group, and not the number of individuals, whenever possible. In cases where the control or treatment group was used repeatedly across multiple comparisons, sample sizes were adjusted by dividing them by the number of times each group was used in the comparison, to account for shared group non-independence. For the cases where the authors reported no significant difference between the two groups (k=2, n=16) but did not provide the values, we used zero as the effect size. To impute the SD for these cases, we used the missing case approach following @nakagawa2022.

### Meta-analysis and meta-regressions

-   main effect model

-   meta-regressions for testing hypothesis

-   meta-regressions for exploratory analyses

Senstivity analysis and Publication Bias Test

### Deviations from registration

**Methods**

**Effect Size Calculation**

To quantify the effect of green nest material on bird fitness, we computed effect sizes using two primary metrics: log response ratio (lnRR) and standardized mean difference with heteroscedasticity correction (SMDH). lnRR was calculated following Hedges et al. (1999) using the `escalc()` function in the R package `metafor` (Viechtbauer, 2010), with `measure=ROM` and `vtype=LS`, the latter ensuring large sample approximation for sampling variance estimation (Bonett, 2008). SMDH was computed similarly using `measure=SMDH`, incorporating bias correction for heteroscedasticity (Bonett, 2009). For studies reporting inferential statistics (e.g., t-values, F-values) but not primary data, we derived SMD estimates using equations from Lajeunesse (2013) and Nakagawa & Cuthill (2007). Effect sizes were adjusted to ensure that positive values corresponded to increased fitness in experimental groups relative to controls. To account for repeated use of control or treatment groups across multiple comparisons, sample sizes were adjusted by dividing them by the number of times each group appeared in comparisons.

**Meta-Analysis and Meta-Regression**

We conducted two separate multilevel meta-analyses (intercept-only models) to estimate the overall effect size using both lnRR and SMDH as dependent variables. Sampling variances were modeled as variance--covariance matrices assuming a 0.5 correlation (ρ) between sampling variances from the same study (Noble et al., 2017). Robust variance estimation was applied using the `robust()` function from `metafor`, employing the CR2 correction from the `clubSandwich` package (Pustejovsky, 2024) for small-sample adjustments. Sensitivity analyses were conducted with correlation values of 0.3 and 0.7, with results reported in supplementary materials.

For moderator analyses, we implemented unimoderator multilevel meta-regressions to investigate heterogeneity sources. These models included key moderators such as parasite type, experimental design, bird species, and trait type. Categorical moderators required a minimum of five data points per level; otherwise, levels were merged or excluded from analyses. To assess heterogeneity, we calculated four metrics (I², τ², CV, and M) following Yang et al. (2023). The proportion of heterogeneity explained by moderators was quantified using R² marginal (Nakagawa & Schielzeth, 2013).

**Publication Bias and Heterogeneity**

@@

@@

**Data Management and Software**

All data extraction and cleaning were conducted in R. The dataset and analysis scripts are publicly available on GitHub (https://github.com/shreyadimri/Green_Nest_Material). Effect size calculations and meta-analyses were performed using the `metafor` package (Viechtbauer, 2010), and additional robust variance estimation was conducted with the `clubSandwich` package (Pustejovsky, 2024).

@@

How many papers in final meta-analysis and how many effect size.

>>>>

pacman::p_load(readr,tidyverse)

@@

```{r load data extraction, include=FALSE, echo=FALSE}

merged_data_extraction <- read_csv("data/03_data_cleaning/merged_data_extraction.csv")

dataset_after_cleaning <- merged_data_extraction %>% filter(fulltext_screening!="excluded")
included_papers_no<-length(unique(dataset_after_cleaning$paper_ID))
```

@@

To obtain any missing data and to locate any grey literature, we wrote to the corresponding author of the articles that passed fulltext screening using a standardized email template (\@figsupp-email). We obtained one unpublished dataset ... \[calculate estimates here\].. Another unpublished dataset was provided by a co-author of this paper. We extracted data for `{r} included_papers_no - 2` published articles and 2 unpublished studies.

@@

```{r load effect size, include=FALSE, echo=FALSE}

inferential_stats<-dataset_after_cleaning%>%filter(statistics_type=="t-test"|statistics_type=="ANCOVA")

contingency_table<-dataset_after_cleaning%>%filter(proxy_decision=="contingency table")

no_ES<-dataset_after_cleaning%>%filter(proxy_comment == "use 0 as ES")

negative_lnRR<-dataset_after_cleaning%>%
  filter(measure_central_tendency_experiment<=0 | measure_central_tendency_control<=0)

We calculated two effect sizes for each pairwise comparision-log response ratio (lnRR) and standardised mean difference with heteroscedasticity correction (SMDH), using `escalc` function (`measure= ROM` and `SMDH` respectively, `vtype= "LS"`) in `metafor` package (@metafor). We specified treatment group as the numerator and the control group as the denominator for lnRR and treatment minus control for SMDH, such that the positive values indicate the trait value increased in the treatment group. In cases such as measures of parasite load or prevalence, laying or hatching dates, courtship time, scab score etc., the sign was inverted to ensure comparability across all fitness proxies. In all cases, the effect size were coded such that the fitness proxies were positively correlated with the overall fitness.

For studies reporting only inferential statistics (e.g., t-tests, ANCOVA), we estimated SMD values (n=`{r} nrow(inferential_stats)`, k=`{r} length(unique(inferential_stats$paper_ID))`) using the equations provided by @thehand2019 . In cases of dichotomous variables (e.g., successful nests, n=`{r} nrow(contingency_table)`, k=`{r} length(unique(contingency_table$paper_ID))`) we used the data to calculate log odds ratio using `escalc` function (`measure=OR`) and convert it to SMD using equations from @thehand2019. Since lnRR cannot be calculated from inferential statistics or dichotomous variables, only SMD under the assumption of homoskedasticity was derived from such data. We excluded these effect sizes along with any zero or negative values (n=`{r} nrow(negative_lnRR)`, k=`{r} length(unique(negative_lnRR$paper_ID))`) when calculating lnRR.

The sample sizes used was the number of nests in each group, and not the number of individuals, whenever possible. In cases where the control or treatment group was used repeatedly across multiple comparisons, sample sizes were adjusted by dividing them by the number of times each group was used in the comparison, to account for shared group non-independence. For the cases where the authors reported no significant difference between the two groups (n=`{r} nrow(no_ES)`, k=`{r} length(unique(no_ES$paper_ID))`) but did not provide the values, we used zero as the effect size. To impute the SD for these cases, we used the missing case approach following @nakagawa2022.

## Meta-analysis

We used rma.mv function from metafor pacjage using 'REML' method to fit all the models. To estimate the overall effect of green nest material addition to the nest on fitness across bird species that perform this behaviour, we fitted two separate multilevel intercept-only meta-analytical using lnRR or SMDH as dependent variables. Sampling variances were modelled as variance-covariance matrices assuming a 0.5 correlation (ρ) between sampling variances from the same study [@noble2017].

*To understand the sources of variation, we calculated Q and total heterogeneity (𝜎2), its source (or relative heterogeneity; I2) and two different metrics to understand its magnitude – a mean-standardized metric (CV) and a variance-mean-standardized metric (M; more in* @yang2025*).*

*To understand the sources of variation or heterogeneity, for the intercept-only models, we will use the pluralistic approach recently suggested by which consists of calculating four metrics of heterogeneity (σ2, I2, CV, M) to comprehensively understand both total and random-effect specific heterogeneity. For the meta-regressions, we will calculate Rmarginal2 to quantify the heterogeneity explained by the moderator(s) in proportion to the total variation (Nakagawa & Schielzeth, 2013).*

## Meta-regression for testing hypothesis

For understanding the effect of green nest material on reproductive success (courtship hypothesis) and on protective effects and health benefits for the nestling (Parental care hypothesis), we ran a multilevel meta-regression with 'hypothesis' (Courtship hypothesis, Parental care hypothesis, Both) as a moderator.

*For meta-regression models, we estimated the percentage of heterogeneity explained by the moderators using R2marginal (Nakagawa & Schielzeth 2013). We graphically presented the results as orchard plots using the R package “orchaRd” v.2.0 (Nakagawa et al. 2023).*

## Meta-regression for exploratory analyses

## Publication Bias Test

@@

## Deviations from registration

@@

@@

We obtained X effect sizes from k studies including How many papers in final meta-analysis and how many effect size.
